---
title: "Practical One"
format: html
---

## Question 1:

### Instructions: 

Find all rows in “airquality” that have missing values. Note that the `airquality` dataset in `R` is always available (just type `airquality` in the console to see it).

### Answer:

```{r}
class(airquality)

airquality_with_missingvalues <<- airquality[
  is.na(airquality$Ozone) | is.na(airquality$Solar.R) | 
    is.na(airquality$Wind) | is.na(airquality$Temp) | 
    is.na(airquality$Month) | is.na(airquality$Day),]

print(airquality_with_missingvalues)

write.csv(airquality_with_missingvalues, file = "_raw_data/airquality_with_missingvalues.csv") #saving our manipulated data 
```

## Question 2:

### Instructions:

Find mean, sd, min, max for each of temperature and ozone level.

### Answer: 

```{r}
ozone_mean <- mean(airquality$Ozone, na.rm = TRUE)
ozone_mean

ozone_sd <- sd(airquality$Ozone, na.rm = TRUE)
ozone_sd

ozone_min <- min(airquality$Ozone, na.rm = TRUE)
ozone_min

ozone_max <- max (airquality$Ozone, na.rm = TRUE)
ozone_max
  
temp_mean <- mean(airquality$Temp)
temp_mean

temp_sd <- sd(airquality$Temp)
temp_sd

temp_min <- min(airquality$Temp)
temp_min

temp_max <- max(airquality$Temp)
temp_max
```

## Question 3:

### Instructions:

For linear regression, parameter estimates can be found as follows. β^=(XTX)−1XTY Here, Y is the response variable, and X is the design matrix. The cars data (an R data set, also always available in `R`) contains two variables: speed and distance to stop. Fit a simple linear regression model to these data, i.e. find the β estimates, using the equation above, and matrix calcuations in R.

### Answer:

```{r}
y_matrix <- cars$dist

x_matrix <- cbind(1, cars$speed)

linear_reg <- function(x, y) {
  
  num_observations <- nrow(x)
  
  num_variables <- ncol(x)

  beta_hat <- solve(t(x) %*% x) %*% t(x) %*% y   

  regression_residuals <- y - x %*% beta_hat

  rss <- t(regression_residuals) %*% regression_residuals  

  regression_variance <- as.numeric(rss) / (num_observations - num_variables)  

  standard_errors <- sqrt(diag(regression_variance * solve(t(x) %*% x)))  

  t_stats <- beta_hat / standard_errors  

  p_vals <- 2 * (1 - pt(abs(t_stats), num_observations - num_variables))

  return(list(
    "Beta Estimates:" = beta_hat, 
    "Estimated Standard Errors:" = standard_errors, 
    "T-Statistics:" = t_stats, 
    "P-Values:" = p_vals
  ))
}

linear_reg(x_matrix, y_matrix)
```

## Question 4:

### Instructions:

Check that you get the same β estimates as when fitting the linear regression model using lm() in R.

### Answers:

```{r}
reg_model <- lm(dist~speed, data = cars)
summary(reg_model)
```

**As we can see from the summary output, the linear regression function created in question 3 gives the same estimates as the "lm()" function.**
